---
title: Survival on the Titanic
author: by XEM Team
output: github_document
---


```{r load-packages-readme, include = FALSE}
library(tidyverse)
library(broom)
library(here)
library(tidymodels)
library(knitr)
library(xaringanthemer)
library(dplyr)
library(vcd)
library(grid)
library(rsample)
library(ggplot2)
library(ggpol)
library(scales)
library(caret)
library(corrplot)
library(ggridges)
library(forcats)
library(scales)
library(skimr)
library(tidypredict)
library(kableExtra)
library(pscl)
```
 
```{r load-data-readme, include=FALSE}
titanic <- read.csv(here::here("data/titanic.csv"))
```


## Summary
The Titanic, said to be “unsinkable”, collided with an iceberg and sank on April 1912, killing 1502 out of 2224 passengers and crew. This was one of the first accidents of this scale with important data about it which was used to then improve maritime safety by passing new policies, better safety procedures and construction to avoid another catastrophe. 

We have decided to analyse original data from the accident and explore how different attributes of passengers are related to their survival in the catastrophe. It must be noted that we have data on 891 passengers out of 2224 people on board and we don't know how representative our sample is of the whole data. These attributes include gender, age and socio-economic status (represented by passenger class). We also looked at a range of other variables including ticket price and port of embarkation. We then moved on to build a logistic regression model to try and predict survival based on these attributes. Lastly, we created a Mosaic plot on the basis of an historic plot from the times of the accident to display the relationship between the explanatory variables and survival. 

#### Age and survival
From the data set, age is given in years. We expect to see a relationship between age and survival based on different age groups rather than the numerical age itself, so we decided to convert continuous ages into categorical groups (to represent children, teenagers, young adults, middle aged adults and elderly people). This allowed us to create a bar plot to show how the age of passengers and their survival are related. We have found that children (0-12 y.o.) indeed have higher survival rates than the other age groups and survival rates generally decrease with age. This can be explained by the fact that it’s human nature to give priority to saving infants and children first, and was indeed reported as orders from the captain.  

#### Gender and survival
During these times, in the spirit of chivalry, women were saved first. This was historically recorded as orders from the captain of the Titanic. Basing on this, we want to verify if gender affected the survival rate. Since this is a historical data, gender is only recorded as male and female, thus we will use binary data for the following research. First, a bar plot is created, using gender as the predictor variable and percentage of survival displayed in terms of colors showing survivals or not as the outcome variable. The graph shows that the percentage of survival of female is 74.2% yet the percentage of survival of male is only 18.9%. We concluded that female have a higher survival rate than male.

#### Wealth and survival
write about 150 words

#### Model
We fitted a linear regression model to our data trying to use the different variables available to us from the data set and compared the predicted properties. We focused on making our model parsimonious and saw that the explanatory variables which allow for the model with better predicted performance are indeed gender, age and passenger class. We then used and introduced different statistical tests (not in the syllabus), as well as a ROC course to display and evaluate the model's strenght on the test data. In short, the model predicts that young, 1st class, female passengers are more likely predicted to survive compared to the counterparts of each of the attributes. It's worthwhile noting that the purpose of a predicitve model when it refers to historic data like this is rather than predict per se, provide an overall better understanding of the data and what it refers to. 

#### Recration of historic plot
```{r gbron-graph-side, echo=FALSE, out.width= "30%", out.extra='style="float:right; padding:10px"'}
include_graphics("https://www.researchgate.net/profile/Michael-Friendly/publication/330916468/figure/fig1/AS:723679168196613@1549549967751/GBrons-chart-of-The-Loss-of-the-Titanic-from-The-Sphere-4-May-1912-Each-subgroup.png")
```
One of the first visualizations (see right) to ever be created with data from the Titanic is a graph by graphic artist G. Bron published on the Sphere (British newspaper) one week after the accident. His work is an early innovation in data display where each subgroup shown by a bar with area proportional to the numbers of cases which today can be seen as an early mosaic plot. We indeed decided to work on our own mosaic plot to display how all the different factors analysed are related to survival in a single visualisation (see below). 
``` {r table-mosaic, echo=FALSE, include = FALSE}
titanic_child <- titanic %>%
  mutate(ischild = ifelse(Age >= 18, "adult", "child"))
  
tbl <- xtabs(~Survived + Pclass + Sex + ischild, titanic_child)
ftable(tbl)
```

``` {r mosaic-plot-readme, echo = FALSE, out.width = "60%", fig.align = 'center'}
mosaic(~ Survived + Pclass + Sex + ischild, data = titanic_child, 
       highlighting = "Survived", highlighting_fill = c("  cornsilk2  ","sandybrown"),
       labeling_args = list(set_varnames = c(Sex = "Gender",
                                            Survived = "Survived",
                                             Pclass = "Passenger Class",
                                             ischild = "Child")),
        set_labels = list(Survived = c("No", "Yes"),
                          Pclass = c("1st", "2nd", "3rd", "Crew"),
                          Sex = c("F", "M"),
                          ischild = c("A","C")),
       main = "Survival on the Titanic" 
       )
```


## Presentation

Our presentation can be found [here](presentation/presentation.html).

## Data 
CKAN training (2014). Titanic [Dataset], viewed 22 October 2021,  <https://data.wu.ac.at/schema/datahub_io/MWViYjhmYTctMzE2NS00OWEwLThmZDgtMTUwZjI4MThiYTJl>.

## References
[1] Source: Friendly, M., Symanzik, J., &amp; Onder, O. (2019, February 6). Royal Statistical Society Publications. Royal Statistical Society. Retrieved November 19, 2021, from https://rss.onlinelibrary.wiley.com/doi/full/10.1111/j.1740-9713.2019.01229.x. 

https://rkabacoff.github.io/datavis/Models.html#Mosaic 

https://www.houstoninjurylawyer.com/titanic-changed-maritime-law/

https://www.datavis.ca/papers/titanic/ 

[2] https://github.com/awesomedata/awesome-public-datasets

[3] https://www.kaggle.com/c/titanic/data
