---
title: "Project proposal"
author: "XEM Team"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(here)

```

## 1. Introduction

Touted as the ultimate in transatlantic travel and said to be “unsinkable”, the Titanic collided with an iceberg on 14 April 1912 on her maiden voyage and sank shortly thereafter on 15 April, killing 1502 out of 2224 passengers and crew. https://rss.onlinelibrary.wiley.com/doi/full/10.1111/j.1740-9713.2019.01229.x We want to see **if and how the chance of survival of Titanic passengers is related to different attributes of the passengers including sex, age, socio-economic status etc.** -- this is the general research question of our project. 

The data set we have used have comes from the Awesome Public Data Sets on GitHub (https://github.com/awesomedata/awesome-public-datasets) and there is no information on where the data originated or how it was collected. However, we have found an article on a similar data set so we have reason to believe that *"the primary sources of data on the Titanic derive from official inquiries launched in Britain and the USA. Shortly after the disaster, the British Parliament authorised the British Board of Trade Inquiry with Lord Mersey as chair. The committee interviewed over 100 witnesses over 36 days of hearings. Their report, issued on 30 July 1912, contained extensive tables of passengers and crew, broken down by age group, gender, class and survival"* https://rss.onlinelibrary.wiley.com/doi/full/10.1111/j.1740-9713.2019.01229.x 

The titanic.csv file contains data for 887 of the real Titanic passengers (2208 total). Each row represents on passenger and there are 12 different columns which describe different attributes about the person including whether they survived, their age, their passenger-class, their sex and the fare they paid. https://web.stanford.edu/class/archive/cs/cs109/cs109.1166/problem12.html

## 2. Data

```{r load-data, message = FALSE}
titanic <- read.csv(here::here("data/titanic.csv"))

glimpse(titanic)
```

Note that we have added the dimensions and codebook for the dataset is in the `README` in the `\data` folder. 

The 12 variables in the data set are:

* `PassangerId`: ID of passanger (from 1 to 891)
* `Survived`: If passenger survived (0 = No, 1 = Yes)
* `Pclass`: Passenger class (1 = 1st, 2 = 2nd, 3 = 3rd) 
  + **note**: this is a proxy for socio-economic status
* `Name`: Name and Surname of passanger, if available
* `Sex`: Gender of passanger (male or female)
  + **note**: this is historical data and the gender of passengers is defined as binary
* `Age`: Age in years (fractional if less than 1, if it's estimated is it in the form of xx.5)
* `SibSp`: Number of siblings/spouses aboard the Titanic
  + Sibling = brother, sister, stepbrother, stepsister
  + Spouse = husband, wife (mistresses and fiancés were ignored)
* `Parch`: Number of parents/children aboard the Titanic
  + **note**: Parent = mother, father
  + **note**: Child = daughter, son, stepdaughter, stepson (some children travelled only with a nanny, therefore parch=0 for them)
* `Ticket`: Ticket number
* `Fare`: Passenger fare (i.e. cost of ticket in USD)
* `Cabin`: Cabin number
* `Embarked`: Port of embarkation (C = Cherbourg, Q = Queenstown, S = Southampton)

https://www.kaggle.com/c/titanic/data


## 3. Data analysis plan
Hypothesis 1: Women have a higher survival rate than men. 
For question 1 we will create a new variable called "survival_rate_gender", using the "survived" column divided by the total number of different gender, as the response variable. We will also use "gender" as predictor variable. 

Hypothesis 2: The younger survivals are more . 
For question 3 we will have to create a new variable called "age range" which groups the ages in to different ranges such 0-15 , 16-30 etc. Since many passengers have unknown ages we will have to remove from the dataset. Then we will visualize the data in to a segmented bar plot using the age range as the response variable and survived as the outcome. 

Hypothesis 3: The higher the ticket price (fare), the higher survival rate. 

Hypothesis 4: If a passenger had a cabin, the higher the possibility to survive. 



## References


