---
title: "Project proposal"
author: "XEM Team"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r load-packages, message = FALSE, echo = FALSE }
library(tidyverse)
library(broom)
library(here)
library(dplyr)
library(ggplot2)
library(ggridges)
library(forcats)
library(scales)
library(skimr)
```

## 1. Introduction

Touted as the ultimate in transatlantic travel and said to be “unsinkable”, the Titanic collided with an iceberg on 14 April 1912 on her maiden voyage and sank shortly thereafter on 15 April, killing 1502 out of 2224 passengers and crew. https://rss.onlinelibrary.wiley.com/doi/full/10.1111/j.1740-9713.2019.01229.x We want to see **if and how the chance of survival of Titanic passengers is related to different attributes of the passengers including sex, age, socio-economic status etc.** -- this is the general research question of our project. 

The data set we have used have comes from the Awesome Public Data Sets on GitHub (https://github.com/awesomedata/awesome-public-datasets) and there is no information on where the data originated or how it was collected. However, we have found an article on a similar data set so we have reason to believe that *"the primary sources of data on the Titanic derive from official inquiries launched in Britain and the USA. Shortly after the disaster, the British Parliament authorised the British Board of Trade Inquiry with Lord Mersey as chair. The committee interviewed over 100 witnesses over 36 days of hearings. Their report, issued on 30 July 1912, contained extensive tables of passengers and crew, broken down by age group, gender, class and survival"* https://rss.onlinelibrary.wiley.com/doi/full/10.1111/j.1740-9713.2019.01229.x 

The titanic.csv file contains data for 887 of the real Titanic passengers (2208 total). Each row represents on passenger and there are 12 different columns which describe different attributes about the person including whether they survived, their age, their passenger-class, their sex and the fare they paid. https://web.stanford.edu/class/archive/cs/cs109/cs109.1166/problem12.html

## 2. Data

```{r load-data, message = FALSE, echo = FALSE}
titanic <- read.csv(here::here("data/titanic.csv"))
titanic <- read.csv(here::here("data/titanic.csv"))
glimpse( titanic )
```

There are 12 variables in the data set.

* **Numerical variables**: discrete: `PassengerID`, `Age`, `SibSp`, `Parch`, `Ticket`, `Cabin`, and continuous: `Fare`.

* **Categorical variables**: ordinal: `Pclass`, and not: `Survived`, `Name`, `Sex`, `Embarked`.

A full description of these variables as well as some notes are all included in the in the `README` in the `\data` folder, which includes the dimensions and codebook for the dataset. The source for these is: https://www.kaggle.com/c/titanic/data.  

## 3. Data analysis plan

```{r , echo=FALSE}
titanic$Survived <- as.factor(titanic$Survived)
titanic$Sex <- as.factor(titanic$Sex)
titanic$Pclass <- as.factor(titanic$Pclass)
```

#### Hypothesis 1: Women have a higher survival rate than men. 

Especially during the times when the sinking of the Titanic accident occurred (1912), in the spirit of chivalry women (and children) were saved first. This was historically recorded as orders from the captain of the Titanic. In order to confirm if this is true and survival rates for women are indeed higher than for men, we will create a bar plot using gender as the predictor variable and frequency displayed in terms of colors showing survivals or not as the outcome variable. Moreover,  we will calculate the survival rate by gender.See the summary statistics below: filtering by the passengers who  survived, we can see that the proportion of women passengers who survived is as high as 0.68 and the proportion of men passengers who survived is almost half of the women's (approximately 0.32).Since the variables that we will be using are categorical, an appropriate statistical method to verify our hypothesis is the Chi-Square test as it will show whether gender and survival are independent or not of one another. 

Important note: the data we are working with is historical data and the gender of passengers is defined as binary. 

```{r gender-survival }
titanic %>%
  ggplot(mapping = aes(x = Sex, fill = Survived)) +
  geom_bar() +
  theme_minimal() +
  scale_fill_viridis_d(name = "Survived", labels = c("No", "Yes" )) +
  labs(x = "Sex", y = "Frequency", title = "Survival Rate by Gender") +
  theme(legend.position = "bottom")
```

```{r survival_rate_gender}
titanic %>%
  filter(Survived == "1") %>%
  count(Sex) %>%
  mutate(prop_survival = n/sum(n))
```

#### Hypothesis 2: The younger survivals are more than the older survivals 
The age of a passenger can be considered to be high priority. We take it for granted that people give priority to saving infants and young children when it comes to providing life saving equipment. Beyond that the survival of  passengers is highly related to their physical abilities which means that the younger the passenger is the higher possibility is to survive.A possible way to visualize the data it is to use a histogram   using "Age" as the predictor variable(X) and "Frequency" as the outcome variable(Y) displayed in terms of colors showing survivals or not. This type of graph will point out the Age with the highest number of survivals. Since the dataset provides a number of values for ages,  we will have to create a new variable called "Age_Range " which groups the ages by scale of 10, that is 0-10, 11-20 etc so that we can find the survival rate for each age range and come out with more concrete conclusions. See the summary statistics below: filtering by the passengers who  survived, we can see that the proportion of  passengers aged between 21-30 years old have the highest probability to survive with 0.246. A statistical method that is useful to answer our hypothesis is to create a box plot  including all the ages so that we can conclude which ages have highest possibility to survive as it displays the median as well as interquartile range, excluding outliers / extreme values. 

```{r age-survival , warning = FALSE}
titanic %>% 
  ggplot(mapping=aes(x = Age, fill = Survived)) +
  geom_histogram(binwidth = 5 ) +
  theme_minimal() +
  scale_fill_viridis_d(name = "Survived", labels = c("No", "Yes")) +
  labs(x = "Age", 
       y = "Frequency", 
       fill = "Survived", 
       title = "Survival rate by age") +
  theme(legend.position = "bottom")
```


```{r age-survival-rate }
titanic %>%
  filter(Survived == "1") %>%
  mutate(Age_Range = case_when(Age >= 0  & Age <= 10 ~ "0-10",
                               Age >= 11 & Age <= 20 ~ "11-20",
                               Age >= 21 & Age <= 30 ~ "21-30",
                               Age >= 31 & Age <= 40 ~ "31-40",
                               Age >= 41 & Age <=50 ~  "41-50",
                               Age >= 51 & Age <=60 ~  "51-60",
                               Age >= 61 & Age <= 70 ~  "61-70",
                               Age >= 71 & Age <= 80 ~ "71-80"))   %>%
  count(Age_Range) %>%
  mutate(prop_survival = n/sum(n)) 
```

#### Hypothesis 3: The higher the class of the passenger, the higher survival rate
The class of the passenger can be considered to be a measure of their socio-economic status: based on our knowledge on the Titanic disaster we expect to find that the higher the class of the passenger, the higher their survival rate. Moreover, we expect the ticket price (`fare` variable) to follow a similar relationship as it is very likely that the ticket price and the class will have a linear relationship. See the summary statistics below: filtering by the passengers who didn't survive, we can see that the proportion  is as high as 0.68 in 3rd class passengers and much lower for first and second class (approximately 0.15 and 0.18 respectively). Based on this, we can also predict that the ticket price will have a positive linear relationship with the  frequency of survivals. A possible graph to display this relationship is the bar plot  having the class to be the predictor variable and the fare displayed in terms of colors showing survivals or not as the outcome variable. This supports our hypothesis as the people who bought an expensive ticket in first class had a higher survival rate than those who bought a cheap  ticket in 2nd and 3rd class. A statistical method to use  in order to support our hypothesis  is to display the regression line on a  scatter plot where fare is the predictor variable and survival rate is the outcome variable to find out  the regression constant so as to  figure out if there is a strong correlation between these two variables. 

```{r fare-survivals-relationship}
 titanic %>%
  ggplot(mapping = aes(x = Pclass, y = Fare, fill = Survived)) +
  geom_bar(stat = "identity", position = "dodge" ) +
  scale_fill_viridis_d(name = "Survived", labels = c("No", "Yes" )) +
  theme_minimal() +
    labs(x = "Class ", y = "Tikcet Price", fill = "Survived", title = "Survival rate by class and fare") +
   theme(legend.position = "bottom") 
```

```{r class-survival}
 titanic %>%
  filter(Survived == "0") %>%
  count(Pclass) %>%
  mutate(prop_death = n/sum(n)) 
```

#### Visualising the data together -- Recreating a G.Bron's historic chart 
Within our project, we would also like to recreate our version of G.Bron's chart of “The Loss of the ‘Titanic'”, from The Sphere, 4 May 1912 -- first know data visualization on the topic. The plot clearly shows the affect of all the different variables we have mentioned before on the survival rate in a single data visualization. We could explore unseen plot types and see which one can recreate similar results, as well as a different visualization we believe better conveys the information. One possible option is a mosaic plot, but we will need to do more research to reach any conclusions.
