---
title: "Project proposal"
author: "XEM Team"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(here)

```

*For instructions on what each section should include, please see the [project page](https://idsed.digital/assessments/project/#proposal) on the course website. Remove this text when completing your proposal*.

## 1. Introduction


## 2. Data

```{r load-data, message = FALSE}
titanic <- read.csv(here::here("data/titanic.csv"))

glimpse(titanic)
```

The titanic.csv file contains data for 891 of the real Titanic passengers. Each row represents one person. The are 12 different columns which describe different attributes about the person including whether they survived, their age, their passenger-class, their sex and the fare they paid. https://web.stanford.edu/class/archive/cs/cs109/cs109.1166/problem12.html  

We have added the dimensions and codebook for the dataset is in the `README` in the `\data` folder. 

## 3. Data analysis plan


## References


