---
title: "Project proposal"
author: "XEM Team"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(here)

```

*For instructions on what each section should include, please see the [project page](https://idsed.digital/assessments/project/#proposal) on the course website. Remove this text when completing your proposal*.

## 1. Introduction
The general research question of the project is  if and how  the death rate of Titanic’s passengers  is related to different attributes of passengers such as sex ,age  ,class etc. We have chosen the dataset “Titanic “ from the website called 'kaggle.com'. The dataset has 891 observations and 12 variables. 

## 2. Data

```{r load-data, message = FALSE}
titanic <- read.csv(here::here("data/titanic.csv"))

glimpse(titanic)
```

The titanic.csv file contains data for 891 of the real Titanic passengers. Each row represents one person. The are 12 different columns which describe different attributes about the person including whether they survived, their age, their passenger-class, their sex and the fare they paid. https://web.stanford.edu/class/archive/cs/cs109/cs109.1166/problem12.html  

We have added the dimensions and codebook for the dataset is in the `README` in the `\data` folder. 

## 3. Data analysis plan
Hypothesis 1: Wemen have a higher survival rate than men. 
For question 1 we will create a new variable called "survival rate", using the "survived" column divided by the total number of different gender, as the response variable. We will also use "gender" as predictor variable. 

Hypothesis 2: The relationship between survival rate and age. 


## References


