---
title: "Who survived the Titanic?"
subtitle: "Factors contributing to the survival of passengers in the historic Titanic accident of 1912 "
author: "XEM Team <br> Chenyu Li, Emma Ghelfi, Maria Stylianou, Xiao Hu"
institute: "University of Edinburgh"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false

---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
titanic <- read.csv(here::here("data/titanic.csv"))
glimpse(titanic)
```

```{r include=FALSE}
style_xaringan(
  title_slide_background_image = "img/titanicpic.jpg"
)
```

class: center, middle

"Touted as the ultimate in transatlantic travel and said to be “unsinkable”, the Titanic collided with an iceberg on 14 April 1912 on her maiden voyage and sank shortly thereafter on 15 April, killing 1502 out of 2224 passengers and crew." [1]

We want to see 
### If and how the chance of survival of Titanic passengers is related to different attributes of the passengers 


---

class: inverse, center, middle

# Section title

---

# Introduction to data set
* The data set comes from the Awesome Public Data Sets on GitHub (https://github.com/awesomedata/awesome-public-datasets) and there is no information on where the data originated or how it was collected.

* There are 12 variables and 891 observations in this data set. 
.pull-left[
- `Name`: Name and Surname of passanger, if available 
- `Survived`: If passenger survived (0 = No, 1 = Yes)
- `Sex`: Gender of passanger (male or female)
- `Age`: Age in years
- `Pclass`: Passenger class (1 = 1st, 2 = 2nd, 3 = 3rd) 
- `Fare`: Passenger fare (i.e. cost of ticket in USD)
- `Embarked`: Port of embarkation (C = Cherbourg, Q = Queenstown, S = Southampton) ]

---
class: inverse, middle, center
# Hypothesis 1: Women have a higher survival rate than men

---


# Introduction
* Especially during the times when the sinking of the Titanic accident occurred (1912), in the spirit of chivalry women (and children) were saved first. This was historically recorded as orders from the captain of the Titanic.

* **geom_bar()** makes the height of the bar proportional to the number of cases in each group. 
---

# Visualization
.pull-left[
```{r survival_rate_gender,  fig.width=4, echo = FALSE}
titanic %>%
  filter(Survived == "1") %>%
  count(Sex) %>%
  mutate(prop_survival = n/sum(n))
```
]
.pull-right[
```{r gender-survival,out.width="100%", fig.width=4, echo=FALSE}
titanic %>%
  ggplot(mapping = aes(x = Sex, fill = Survived)) +
  geom_bar() +
  theme_minimal() +
  scale_fill_viridis_d(name = "Survived", labels = c("No", "Yes" )) +
  labs(x = "Sex", 
       y = "Frequency", 
       title = "Survival Rate by Gender") +
  theme(legend.position = "bottom")
```

]


---
class: inverse, middle, center

# Using xaringan

---






---
# Embarked/Class/Fare 

The class of the passenger can be considered to be a measure of their socio-economic status: based on our knowledge on the Titanic disaster we expect to find that the higher the class of the passenger, the higher their survival rate. 

```{r class-survived ,echo=FALSE}
titanic$Survived <- as.factor(titanic$Survived)
titanic$Pclass <- as.factor(titanic$Pclass) 

titanic_class <- titanic %>%
  group_by( Pclass ) %>%
 count(Survived) %>%
 mutate(pct = n/sum(n) * 100)

titanic_class %>%
ggplot(aes(x=Pclass , y = pct ,  fill = Survived)) +
  geom_bar( position = position_stack(), stat = "identity", width = .7 ) +
  geom_text(aes(y = pct, label = pct) ,position = position_stack(vjust = 0.5), size = 3.5 )
```


---

class: inverse, center, middle

# Modelling

---

class: inverse, center, middle

# Recreating G. Bron's "Sphere" Diagram

---
.pull-left[
* The Sphere was a popular British illustrated weekly newspaper, published by the Illustrated London News Group from January 1900 until June 1964. It was dedicated to worldwide reporting on popular issues. 
* 4 May 1912, three weeks after the Titanic disaster
* it published a chart (Figure 1, page 16) by the graphic artist G. Bron using data released the week before by the House of Commons.
]


.pull-right[
```{r Gbron-original-chart, echo = FALSE, out.width = "70%", fig.align = "right", fig.cap = "G.Bron's chart of 'The Loss of the Titanic', from The Sphere, 4 May 1912." }
include_graphics("https://www.researchgate.net/profile/Michael-Friendly/publication/330916468/figure/fig1/AS:723679168196613@1549549967751/GBrons-chart-of-The-Loss-of-the-Titanic-from-The-Sphere-4-May-1912-Each-subgroup.png")
```
]



---

#Resources
[1] Source: Friendly, M., Symanzik, J., &amp; Onder, O. (2019, February 6). Royal Statistical Society Publications. Royal Statistical Society. Retrieved November 19, 2021, from https://rss.onlinelibrary.wiley.com/doi/full/10.1111/j.1740-9713.2019.01229.x. 
