---
title: "Who survived the Titanic?"
subtitle: "Factors contributing to the survival of passengers in the historic Titanic accident of 1912 "
author: "XEM Team <br> Chenyu Li, Emma Ghelfi, Maria Stylianou, Xiao Hu"
institute: "University of Edinburgh"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false

---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
titanic <- read.csv(here::here("data/titanic.csv"))
glimpse(titanic)
```

```{r include=FALSE}
style_xaringan(
  title_slide_background_image = "img/titanicpic.jpg"
)
```

class: center, middle



"Touted as the ultimate in transatlantic travel and said to be “unsinkable”, the Titanic collided with an iceberg on 14 April 1912 on her maiden voyage and sank shortly thereafter on 15 April, killing 1502 out of 2224 passengers and crew." [1]

We want to see 
### If and how the chance of survival of Titanic passengers is related to different attributes of the passengers 


---

class: inverse, center, middle

# Data Set

---



---

class: inverse, center, middle

# Modelling

---

class: inverse, center, middle

# Recreating G. Bron's "Sphere" Diagram

---

``` {r Gbron-caption, echo = FALSE}
gbron_cap <- "G.Bron's chart of 'The Loss of the Titanic', \\n
from The Sphere, 4 May 1912."
```

```{r Gbron-original-chart, echo = FALSE, out.width = "30%", fig.align = "right", fig.cap = gbron_cap }
include_graphics("https://www.researchgate.net/profile/Michael-Friendly/publication/330916468/figure/fig1/AS:723679168196613@1549549967751/GBrons-chart-of-The-Loss-of-the-Titanic-from-The-Sphere-4-May-1912-Each-subgroup.png")
```

* The Sphere was a popular British illustrated weekly newspaper, published by the Illustrated London News Group from January 1900 until June 1964. It was dedicated to worldwide reporting on popular issues. On 4 May 1912, only three weeks after the Titanic disaster, it published a chart (Figure 1, page 16) by the graphic artist G. Bron using data released the week before by the House of Commons.
---
#Resources
[1] Source: Friendly, M., Symanzik, J., &amp; Onder, O. (2019, February 6). Royal Statistical Society Publications. Royal Statistical Society. Retrieved November 19, 2021, from https://rss.onlinelibrary.wiley.com/doi/full/10.1111/j.1740-9713.2019.01229.x. 
