---
title: "Who survived the Titanic?"
subtitle: "Factors contributing to the survival of passengers in the historic Titanic accident of 1912 "
author: "XEM Team <br> Chenyu Li, Emma Ghelfi, Maria Stylianou, Xiao Hu"
institute: "University of Edinburgh"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false

---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
titanic <- read.csv(here::here("data/titanic.csv"))
glimpse(titanic)
```

```{r include=FALSE}
style_xaringan(
  title_slide_background_image = "img/titanicpic.jpg"
)
```


class: center, middle

"Touted as the ultimate in transatlantic travel and said to be “unsinkable”, the Titanic collided with an iceberg on 14 April 1912 on her maiden voyage and sank shortly thereafter on 15 April, killing 1502 out of 2224 passengers and crew." [1]

We want to see 
### If and how the chance of survival of Titanic passengers is related to different attributes of the passengers 


---

class: inverse, center, middle

# Section title

---

# Introduction to data set
* The data set comes from the Awesome Public Data Sets on GitHub (https://github.com/awesomedata/awesome-public-datasets) and there is no information on where the data originated or how it was collected.

* There are 12 variables and 891 observations in this data set. 
.pull-left[
- `Name`: Name and Surname of passanger, if available 
- `Survived`: If passenger survived (0 = No, 1 = Yes)
- `Sex`: Gender of passanger (male or female)
- `Age`: Age in years
- `Pclass`: Passenger class (1 = 1st, 2 = 2nd, 3 = 3rd) 
- `Fare`: Passenger fare (i.e. cost of ticket in USD)
- `Embarked`: Port of embarkation (C = Cherbourg, Q = Queenstown, S = Southampton) ]

---
class: inverse, middle, center
# Hypothesis 1: Women have a higher survival rate than men

---


# Introduction
* Especially during the times when the sinking of the Titanic accident occurred (1912), in the spirit of chivalry women (and children) were saved first. This was historically recorded as orders from the captain of the Titanic.

* **geom_bar()** makes the height of the bar proportional to the number of cases in each group. 
---

# Visualization
.pull-left[
- Use gender as the predictor variable and frequency displayed in terms of colors showing survivals or not as the outcome variable
]

.pull-right[
```{r gender-survival,out.width="100%", fig.width=4, echo=FALSE}
titanic$Survived <- as.factor(titanic$Survived)
titanic$Pclass <- as.factor(titanic$Pclass) 

titanic %>%
  ggplot(mapping = aes(x = Sex, fill = Survived)) +
  geom_bar() +
  theme_minimal() +
  scale_fill_viridis_d(name = "Survived", labels = c("No", "Yes" )) +
  labs(x = "Sex", 
       y = "Frequency", 
       title = "Survival Rate by Gender") +
  theme(legend.position = "bottom")
```
]

=======




---
# Summary
.pull-left[
- The proportion of women passengers who survived is 0.68 
- The proportion of men passengers who survived is 0.32
]
.pull-right[
```{r survival_rate_gender,  fig.width=4, echo = FALSE}
titanic %>%
  filter(Survived == "1") %>%
  count(Sex) %>%
  mutate(prop_survival = n/sum(n))
```
]
---
class: inverse, middle, center

# Using xaringan

---





---
class: inverse, middle, center
# Hypothesis 2: The younger survivals are more than the older survivals 
---
The age of a passenger can be considered to be high priority. We take it for granted that people give priority to saving infants and young child when it comes to providing life saving equipment. Beyond that the survival of passengers is highly related to their physical abilities which means that the younger the passenger is the higher possibility is to survive. 
---
# visualization
.pull-right[
- Use "Age" as the predictor variable(X) and "Frequency" as the outcome variable(Y) displayed in terms of colors showing survivals or not.
- Create a a new variable called "Age_Range" which groups the ages by scale of 10 so that we can find the survival rate for each age range and come outwith more concrete conclusions.
]

.pull-left[
```{r age-survival,out.width="100%", fig.width=4, echo=FALSE}
titanic %>% 
  ggplot(mapping = aes(x = Age, fill = Survived)) +
  geom_histogram(binwidth = 6 ) +
  theme_minimal() +
  scale_fill_viridis_d(name = "Survived", labels = c("No", "Yes")) +
  labs(x = "Age", 
       y = "Frequency", 
       fill = "Survived", 
       title = "Survival rate by age") +
  theme(legend.position = "bottom")
```
]

---

# Summary
The proportion of passengers aged between 21-30 years old have the highest probability to survive with 0.246.

```{r age-survival-rate,fig.width=6,echo=FALSE}
titanic %>%
  filter(Survived == "1") %>%
  mutate(Age_Range = case_when(Age >= 0  & Age <= 10 ~ "0-10",
                               Age >= 11 & Age <= 20 ~ "11-20",
                               Age >= 21 & Age <= 30 ~ "21-30",
                               Age >= 31 & Age <= 40 ~ "31-40",
                               Age >= 41 & Age <=50 ~  "41-50",
                               Age >= 51 & Age <=60 ~  "51-60",
                               Age >= 61 & Age <= 70 ~  "61-70",
                               Age >= 71 & Age <= 80 ~ "71-80")) %>%
  count(Age_Range) %>%
  mutate(prop_survival = n/sum(n)) 
```

---
# Embarked/Class/Fare 

The class of the passenger can be considered to be a measure of their socio-economic status: based on our knowledge on the Titanic disaster we expect to find that the higher the class of the passenger, the higher their survival rate.```{r class-survived ,echo=FALSE}
titanic$Survived <- as.factor(titanic$Survived)
titanic$Pclass <- as.factor(titanic$Pclass) 

titanic_class <- titanic %>%
  group_by( Pclass ) %>%
 count(Survived) %>%
 mutate(pct = n/sum(n) * 100)

titanic_class %>%
ggplot(aes(x=Pclass , y = pct ,  fill = Survived)) +
  geom_bar( position = position_stack(), stat = "identity", width = .7 ) +
  geom_text(aes(y = pct, label = pct) ,position = position_stack(vjust = 0.5), size = 3.5 )
```


---

class: inverse, center, middle

# Modelling

---

class: inverse, center, middle

# Recreating G. Bron's "Sphere" Diagram

---
## G. Bron's original Diagram on "the Sphere", 1912
.left-column[
```{r Gbron-original-chart, echo = FALSE, out.width = "120%", fig.align = "right", fig.cap = "G.Bron's chart of 'The Loss of the Titanic', 4 May 1912." }
include_graphics("https://www.researchgate.net/profile/Michael-Friendly/publication/330916468/figure/fig1/AS:723679168196613@1549549967751/GBrons-chart-of-The-Loss-of-the-Titanic-from-The-Sphere-4-May-1912-Each-subgroup.png")
```
]
.right-column[
+ The Sphere: popular British illustrated weekly newspaper (1900-1964). It was dedicated to worldwide reporting on popular issues. 
{{content}}

]

--

+ G. Bron, graphic artist, used data released the week before by the House of Commons for the visualisation pubblished on 4 May 1912, **three weeks after the Titanic disaster**
{{content}}

--

+  Bron’s graph is an early innovation in data display:
{{content}}


--

    + each subgroup is shown by a bar whose area is proportional to the numbers of cases
{{content}}   

--

    + Within the passenger classes, the bars are subdivided by gender for adults, while children are shown as a separate group. 
{{content}}

--

    + It also includes summary panels, showing the totals for passengers and crew.
{{content}}
  
--

+  Today, we might describe this as an early form of a mosaic plot, or as an area-proportional back-to-back array of bar charts





---
```{r }
#titanic_freq <- titanic %>%
#  mutate(ischild = ifelse(Age >= 18, "adult", "child"),
#         issurvived = ifelse(Survived == 0, "no", "yes")
#        ) %>%
#  group_by(ischild, Sex, Pclass, issurvived) %>%
#  summarise(Freq = n()) %>%

#vcd::doubledecker(issurvived ~ Pclass + Sex, 
#                  xtabs(Freq ~ Sex + issurvived + Pclass, titanic_freq))
```
---
#Resources
[1] Source: Friendly, M., Symanzik, J., &amp; Onder, O. (2019, February 6). Royal Statistical Society Publications. Royal Statistical Society. Retrieved November 19, 2021, from https://rss.onlinelibrary.wiley.com/doi/full/10.1111/j.1740-9713.2019.01229.x.