---
title: "Who survived the Titanic?"
subtitle: "Factors contributing to the survival of passengers in the historic Titanic accident of 1912 "
author: "XEM Team <br> Chenyu Li, Emma Ghelfi, Maria Stylianou, Xiao Hu"
institute: "University of Edinburgh"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false

---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
titanic <- read.csv(here::here("data/titanic.csv"))
glimpse(titanic)
```

```{r include=FALSE}
style_xaringan(
  title_slide_background_image = "img/titanicpic.jpg"
)
```

class: center, middle

## A statement of the overall goal / research question

Touted as the ultimate in transatlantic travel and said to be “unsinkable”, the Titanic collided with an iceberg on 14 April 1912 on her maiden voyage and sank shortly thereafter on 15 April, killing 1502 out of 2224 passengers and crew.
We want to see **if and how the chance of survival of Titanic passengers is related to different attributes of the passengers including sex, age, socio-economic status etc.** 

---

class: inverse, center, middle

# Section title

---

# Introduction to data set
* The data set comes from the Awesome Public Data Sets on GitHub (https://github.com/awesomedata/awesome-public-datasets) and there is no information on where the data originated or how it was collected.

* There are 12 variables and 891 observations in this data set. 
.pull-left[
* **Name:** Name and Surname of passanger, if available 
* **Survived:** If passenger survived (0 = No, 1 = Yes)
* **Sex:** Gender of passanger (male or female)
* **Age:** Age in years
* **Pclass:** Passenger class (1 = 1st, 2 = 2nd, 3 = 3rd) 
* **Fare:** Passenger fare (i.e. cost of ticket in USD)
* **Embarked:** Port of embarkation (C = Cherbourg, Q = Queenstown, S = Southampton) ]

---
class: inverse, middle, center
# Hypothesis 1: Women have a higher survival rate than men

---


# Introduction
* Especially during the times when the sinking of the Titanic accident occurred (1912), in the spirit of chivalry women (and children) were saved first. This was historically recorded as orders from the captain of the Titanic.

* **geom_bar()** makes the height of the bar proportional to the number of cases in each group. 
---

# Visualization
.pull-left[
- Some text
- goes here
]
.pull-right[
```{r gender-survival, warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
titanic %>%
  ggplot(mapping = aes(x = Sex, fill = Survived)) +
  geom_bar() +
  theme_minimal() +
  scale_fill_viridis_d(name = "Survived", labels = c("No", "Yes" )) +
  labs(x = "Sex", 
       y = "Frequency", 
       title = "Survival Rate by Gender") +
  theme(legend.position = "bottom")
```
]
---


class: inverse, middle, center

# Using xaringan

---

# xaringan

- The presentation is created using the `xaringan` package

- Use `---` to separate slides and `--` for incremental builds

--

- Like this

---

# Layouts

You can use plain text

- or bullet points

.pull-left[
or text in two columns $^*$
]
.pull-right[
- like
- this
]

.footnote[
[*] And add footnotes
]

---

# Code

```{r boring-regression}
# a boring regression
model <- lm(dist ~ speed, data = cars)
tidy(model)
glance(model)
```

---

# Plots

```{r recode-species, echo = FALSE}
# In this chunk I'm doing a bunch of analysis that I don't want to present 
# in my slides. But I need the resulting data frame for a plot I want to present.
iris_modified <- iris %>%
  mutate(Species = fct_other(Species, keep = "setosa"))
```

```{r plot-iris, echo = FALSE}
# Code hidden with echo = FALSE
# Uses modified iris dataset from previous chunk
# Play around with height and width until you're happy with the look
ggplot(data = iris_modified, mapping = aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
  geom_point() + 
  theme_minimal() # theme options: https://ggplot2.tidyverse.org/reference/ggtheme.html
```

---

## Plot and text

.pull-left[
- Some text
- goes here
]
.pull-right[
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
ggplot(penguins, aes(x = bill_length_mm, y = species, color = species)) +
  geom_boxplot() +
  theme_minimal()
```
]

---

# Tables

If you want to generate a table, make sure it is in the HTML format (instead of Markdown or other formats), e.g.,

```{r iris-table, echo = FALSE}
kable(head(iris), format = "html")
```

---

# Images

```{r castle, echo = FALSE, out.width = "60%", fig.align = "center", fig.cap = "Image credit: Photo by Jörg Angeli on Unsplash."}
include_graphics("https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80")
```

Or you can also include a full page image. See next slide.

---

background-image: url(https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80)

---

# Math Expressions

You can write LaTeX math expressions inside a pair of dollar signs, e.g. &#36;\alpha+\beta$ renders $\alpha+\beta$. You can use the display style with double dollar signs:

```
$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$
```

$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$

Limitations:

1. The source code of a LaTeX math expression must be in one line, unless it is inside a pair of double dollar signs, in which case the starting `$$` must appear in the very beginning of a line, followed immediately by a non-space character, and the ending `$$` must be at the end of a line, led by a non-space character;

1. There should not be spaces after the opening `$` or before the closing `$`.

1. Math does not work on the title slide (see [#61](https://github.com/yihui/xaringan/issues/61) for a workaround).

---

# Feeling adventurous?

- Want to find out more about `xaringan`? See https://slides.yihui.name/xaringan/#1.

- You are welcomed to use the default styling of the slides. In fact, that's what I expect majority of you will do. You will differentiate yourself with the content of your presentation.

- But some of you might want to play around with slide styling. The 
`xaringanthemer` provides some solutions for this that: https://pkg.garrickadenbuie.com/xaringanthemer.

- And if you want more bells and whistles, there is also `xaringanExtra`: https://pkg.garrickadenbuie.com/xaringanExtra.
